\section*{Gratitude}
\label{chap:gratitude}

My gratitude goes out to my support network from my university, my family, and my friends.
Especially my supervisor Marc Schickler for offering encouragement and praise when things worked -- I appreciate the support.
He also offered up excellent points for improving my thesis in his corrections.
My extended family for the many nice questions about what I was doing and sitting through the long explanations with sufficient interest to keep me going.
Maybe someday I will find a single sentence that sufficiently explains everything comprehensibly for "non-computer" people.
My aunt Sibille and my parents Katja and Stefan for their hard work of proof reading this thesis: all remaining errors are my own.
Furthermore my friends for regularly asking how I was doing, thus building up the pressure to keep working continuously.

This work would not have been possible without the great tools and libraries and the many people that implemented and documented them.
Therefore my heartfelt thanks go out first and foremost to the open source community that made building this thesis a possibility: Tox for providing a peer to peer fully encrypted communication channel and Golang for the access to source code when things didn't work as expected.
Worthy of special mention is the Github user Codedust for maintaining the Golang Tox wrapper on which I built this thesis.
Special thanks are due for their support whenever I had problems or feature requests and always received help and support.
