\chapter{Introduction}
\label{chap:Introduction}

In this chapter we first give a short overview over the thesis.
Then we discuss the motivation for our work and introduce the name given to it.
Next we place the master thesis in its context.
Finally we briefly summarize the structure of this document.

\section{Overview}

We propose the implementation of a file synchronization program in Golang via the Tox protocol.
Thus we first explain our motivation for this thesis and what features we would like the implementation to fulfill.
Then we will expand on the specification and define the scope of this work in preparation.
Apart from the theoretical work a large part of this thesis is also the implementation of a proof of concept.
Therefore we will also expand on problems and solutions we encountered while working on the software aspect.
Finally we will look back at the implementation and compare it to the theoretical ground work.
For better comparison a brief discussion on similarities and differences to existing file synchronization services will also be included.

\section{Motivation}

The thesis is mainly motivated by the revelations of drag net surveillance by Edward Snowden and the compliance by so called trusted service providers beyond even their legal obligations.
Thanks to his sacrifice a global conspiracy of governmental agencies that do massive online surveillance of most internet communications was brought into the public light.
Since the majority of the global internet players is based in the United States of America, this had huge implications for all manner of information exchanged.
Many services and programs used by the majority of internet users has been shown to be either at risk of being compromised.
This includes internet giants such as Facebook and and Google~\cite{web:site:tg:internet_giants}.
Even Microsoft, producer of the most widely used operating system for personal computers, complied~\cite{web:site:tg:microsoft}.

While public outrage lasted just as long as the news reported on it, the revelations had a major impact in the technical community.
New software solutions are required to enable users to take back their privacy on the internet without sacrificing usability and ease of access.
One example is the Open Whisper Systems group.
Their stated goal on the website is "[...] we're working to advance the state of the art for secure communication, while simultaneously making it easy for everyone to use."~\cite{web:site:whispersystems}.
Another example is the Tox instant messenger community that is building a free, open source, peer to peer Skype alternative~\cite{web:site:tox}.

The main goal of this thesis is to tackle the problem of cloud data storage, often used for distributed file access across multiple computers.
Existing examples of this include Google Drive~\cite{web:site:gdrive} and Dropbox~\cite{web:site:dropbox}.
While the data transfered is now encrypted in transit in most cases, it is not encrypted while residing on the servers.
This allows anyone with access to the server to fully access the data, so long as it was not already previously encrypted by the user.
Before Snowden the third parties were entrusted with the safe keeping of the stored data under the presumption that no unauthorized access was allowed or even possible.
However this trust has been lost now that the public has been made aware of the possibilities of state organizations to access any data without due legal process.

One solution to this problem is encrypting everything client side before uploading the data to third parties.
An example of a service that does this is Boxcryptor~\cite{web:site:boxcryptor}.
However this principle requires active involvement from the user and is thus an extra hurdle in securing one's data.
Perhaps simply forgoing a third party would be a possible solution.
Examples of existing peer to peer solutions are BitTorrent Sync~\cite{web:site:bittorrent_sync} and Syncthing~\cite{web:site:synthing}.
However now the main advantage of a third party is lost: namely the ability to store data in a way that it is readily accessible even if every private peer is currently offline.
Syncthing notably explicitly states that the user can enable access for trusted third parties, but as it stands now this third party would receive an unencrypted access to any data stored there.

Thus our motivation is to create a proof of concept of a file synchronization software suite.
Stated goals are peer to peer communications, bypassing the requirement of a centrally hosted third party service.
To harden such a peer to peer network it is necessary to encrypt all communication between the peers.
This is not only required to hinder Wifi sniffing when using unsecured wireless access: Snowden also revealed that most data passing through internet exchange points is surveiled~\cite{web:site:heise:decix}.
Instead of integrating and mixing encrypted communications with the file synchronization commands required for such software to work, we utilize an existing secure peer to peer communication infrastructure in the form of Tox, on top of which we will propose a protocol for the sole task of file synchronization.

To differentiate our proof of concept implementation from existing solutions a unique name was required.
We chose the rock-forming mineral tinzenite for the name.
Specifically, Tinzenite is the name of the peer network, which in turn is compromised of two distinct software peers which are based on a common protocol and communication standard.

\section{Project Context}

This paper is the master thesis of Tamino P.S.M. Hartmann, written at the Faculty of Engineering, Computer Science, and Psychology~\cite{web:site:faculty} at the Institute of Databases and Information Systems~\cite{web:site:institute} at the University of Ulm~\cite{web:site:uni_ulm}, Germany.
The supervisor was Marc Schickler and the examiners were Professor Doctor Manfred Reichert and Professor Doctor Martin Theobald.
The proposed work is to create and implement a proof of concept for a fully encrypted peer to peer file synchronization software.
Stated goals are the study of the feasibility of such a work and the exploration of associated technologies.
These specifically include Tox, Golang, and Hadoop.
